Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-11 how.tex
Updates from ShareLaTeX

\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 %\usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage[running]{lineno}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath}  
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=right,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, Line,bPruning,clister,train,leafs,weightedFeatures, exemplar,nearestLine,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}

\title{Learning Actionable Analytics 
 (with applications for reducing defects and reducing runtimes)}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
Rahul Krishna, Tim Menzies, Xipeng Shen\\
        Computer Science, NcState, USA\\
       \{i.m.ralk, tim.menzies, xipengshen\}\@gmail.com
% 3rd. author
 % use '\and' if you need 'another row' of author names
% 4th. author
\and
 Andrian Marcus\\
       Computer Science, UtDallas, USA \\
       amarcus\@utdallas.edu }

@@

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, the improvements generated by HERE can reduce
 defect counts as well as runtimes by up to
 50\% of their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Defect prediction, configuration, prediction, planning, instance-based reasoning.
  \end{IEEEkeywords}

\section{Introduction}
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself. For example:
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In this paper, we say   a ``plan'' is a  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every $Z$  that requests a plan $\delta$, we want
to adjust the attributes $Z_j$ as follows:
\begin{equation}\label{eq:one}
Z_j = Z_j + \delta_j
\end{equation}
The usual way such plans are generated in the SE literature
is via some search-based method. Recently, we have made  some progress
in this area with the  GALE optimizer~\cite{krall14}. 
Like other search-based algorithms,
 GALE executes by proposing multiple $\delta$s
 (one for each generation in an evolutionary program).
 In practice, GALE makes
 extensive changes to initial sample of project data
 (perhaps 
100s to 100,000s of mutations). Each of these mutations
is reassessed using some domain model.
GALE runs faster than standard search-based SE tools~\cite{deb00a,zit02} since
it explores  the principle components
 of the project data (rather than   the raw data). Since it explores
 a smaller space, it also
 terminates much faster than standard optimiser; e.g. for one large model, 4 minutes versus 7 hours~\cite{krall14}.

When presented to some  users, the GALE approach was challenged by
two business users.   Note that the following issues
challenge not just GALE, but also  any  optimization algorithm  that makes
extensive use of mutation  and domain models such as
NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}.

One  user asked:
\begin{quote}
ISSUE~\#1: {\em ``What if I only want the minimum useful change, not
the maximum possible?''}
\end{quote}
Note here that this business user is asking for solutions  that do not require extensive 
changes to their current  approaches.

@@

Every challenge is an opportunity to explore  a new approach.
If we  restrict $\delta$ to very small changes, then any 
predictive data miner applied to the original data could also be used
to assessed the mutated data (proviso: we have to guarantee that
the mutations do not leave the space of examples present in the original data).
Note that this approach addresses ISSUE~\#1 since it only uses minimal changes.
It also addresses  ISSUE~\#2 since, when lack a domain model, we automatically
create one (via data mining).

This paper presents HOW, a novel planner for learning actionable analytics via
minimal changes. 
In the experiments shown below, we see that HOW's plans  can significantly
reduce the  expected
value of:
\bi
\item The software defects found in  the object-oriented JAVA systems explored in 2010 by Jureczko et al.~\cite{jureczko10};
\item The runtime of the  software systems whose configurations
are explored in 2013 by Seigmund et al.~\cite{sven12};
\ei
This paper makes five contributions:
\be
\item The HOW  instance-based planner (which has not been presented before).
\item The experiments of this paper show that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\item Apart from HOW, we  also offers a general framework for using  predictive system to build planners;
\item Further, we offer a  test to recognizes  when we should {\em not}  plan: specifically,
planning is
not recommended in domains with  poorly performing predictors;
\item Last, we  offer a general verification rig for assessing the value of plans generated within this rig.
\ee

\section{Related Work}
\subsection{Accessing a Trustworthy model}
The premise of this work is that we do {\em not} have
a ready-to-use

In some domains, there is ready access to 
trustworthy domain model that can offer an assessment
of newly generated examples. For example, in other work
we have used executing source code as the ``model'' to check
if our mutations to test suites
  minimize
test suite size while maximizing the number of statements covered~\cite{}

For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}). 
Also, even when there is an existing model, they can require
constant  maintenance lest they become out-dated. Elsewhere, we have described our
extensions to the USC models to enable reasoning about agile software developments. 
It took many months to implement and certify those extensions~\cite{me09i,me09j}.
The problem of model maintenance is another
motivation to look for alternate methods that can be automatically updated whenever new data arrives.
Further, sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

@@

\subsection{Design}
A general model for {\em predictors} is as follows. 
During testing, each test instance gains a prediction 
by finding which region of the data it falls into.
During training, those regions are inferred from
divisions that ``chunk'' together regions with instances
with similar properties. These dividers can be categorical (as in the case of decision trees) or probabilistic (as in the
case of Naive Bayes) or learned via some distance minimization
process 
(as in the case of support vector machines) or generated
directly by some clustering algorithm.

@@

Data can be rapidly and recursively divided   as follows.
Find   two   distance instances,  $X,Y$
by picking any instance $W$ at random, then setting $X$ to its most
distant instance, then setting $Y$ to the instance most distant from
$X$~\cite{fastmap}
(this requires only $O(2N)$ comparisons
of $N$ instances).
Next, project each instance $Z$
onto a line that  runs between $X,Y$ using the cosine
rule of \fig{where}b. After that,  split the data at the median $x$ value of all instances and
recurses on each half  (stopping when
one half has less  than $\sqrt{N}$ of the original population). For more details. \fig{howcode}, line~76,

@@

In the \fig{where}a, to compute distances, we use
the Euclidean measure recommended for
instance-based reasoning by Aha et al.~\cite{aha91};
i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
where values are  normalized 0..1 for the range min..max and 
$w_i$ is some importance weight given to attribute $i$.
Usually, $w_i=1$ but it can be adjusted using, say,
the feature weighting methods of \fig{where}d.

@@

Let   $X,Y$ be two points joined by  a line  length $c$.
A third point, $Z$, has distances  $a=dist(Z,X)$ and
$b=dist(Z,Y)$ to $X,Y$, respectively.
According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
Further, according to Pythagoras' theorem, $Z$ stands at a distance
$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. For more details, see  \fig{howcode}, line~59.

@@

HERE's feature weighting algorithm
comes from the CART regression tree learner~\cite{Breiman1984}.
It sorts independent variables
 according to how well they can reduce the variability
of a  numeric objective.
If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
then the expected
value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
where $v$ is the standard deviation of the objective.
A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
then split at $\argmin_i w_i$, then recursing into each half.  For more details, see  \fig{howcode}, line~109.

@@

\input{_tex/code}

Those same regions can be used for instance-based 
{\em planning}. However,  instead of leaving test instances in the regions
found during training, we {\em displace} them towards regions
with better objective scores.
To achieve this, 
during training, HOW clusters data using the procedure of \fig{where}a,
then draws {\tt Line}s between the centroids of  clusters $X,Y$ that are nearest
neighbors.
By convention, we say that the cluster containing $X$ has the  better objective scores. This means we can say that the {\tt Line} is a useful direction for displacing
data; i.e.  away from $Y$ and towards $X$ (for an implementation of such a {\tt Line} class, see \fig{howcode}, line~1).
 During testing,  HOW   finds the  nearest {\tt Line} to every test instance. That  test instance is   then displaced towards the better end
of that line; i.e. towards  $X$ (see \fig{howcode}, line 53).

An important design decision with HOW 
is how to determine the exact {\em direction} and {\em magnitude} of the
displacement.  Following on from the discussion around   ISSUE~\#1,
we note that a design goal of HOW is
{\em minimal displacements} of project data.
To implement minimal displacement:
\bi
\item
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   $B \in \{25,50,100\}\%$.
\item
HOW also  limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.
\ei
Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test instance.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 19. In summary, 
If  $Z_j$ is an independent variable within the
test instance $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above. Note that this
this equations satisfies our definition of a ``plan'' as offered in
\eq{one} in the {\em Introduction}

The experiments of this work showed that different kinds of data require
different kinds of exemplars.
\bi
\item
For the Jureczko data, the objective score
is a boolean that is true if defects have been noted
for that class. 
For this kind of data with discrete objectives,
the data, 
has very little to  differentiate between the objective scores
so we just set the exemplar to  
an instance synthesized from the median values of the instances
in that cluster.
\item
For the Seigmund data,
the objective score is the runtime of the test suite of a compiled program.
For this kind of data with numeric objectives,
the exemplar  is the instance in a cluster
with  best objective score.
\ei

@@

  The Seigmund data sets are far more varied than the Jurecko data (since they contain every configuration option in the Makefiles
  for BerkeleyDB-- the Java  and  Windows  versions,  Apache,  SQLite,  LLVM,and x264).
  \fig{cpm} lists the size of Seigmund attribute sets. For the Siegmind data, 
  25 times, 
  we randomized the order of the data, trained on the first half (using
  Random Forest Regressor), then tested on the second half.

>>>>

\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 %\usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage[running]{lineno}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath}  
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max} 
%\usepackage[usenames]{xcolor}

@@

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=right,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}

\title{Learning Actionable Analytics 
 (with applications for reducing defects and reducing runtimes)}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each Slope of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
Rahul Krishna, Tim Menzies, Xipeng Shen\\
        Computer Science, NcState, USA\\
       \{i.m.ralk, tim.menzies, xipengshen\}\@gmail.com
% 3rd. author
 % use '\and' if you need 'another row' of author names
% 4th. author
\and
 Andrian Marcus\\
       Computer Science, UtDallas, USA \\
       amarcus\@utdallas.edu }

@@

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, those improvements   reduce
 defect counts and  runtimes by   
 (25,33)\% (median)   to (37,78)\% (max) of
 their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Defect prediction, configuration, prediction, planning, case-based reasoning.
  \end{IEEEkeywords}

\section{Introduction}
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself. For example:
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In this paper, we say   a ``plan'' is a  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every $Z$  that requests a plan $\delta$ to
  adjust   attributes $Z_j$:
\begin{equation}\label{eq:one}
Z_j = Z_j + \delta_j
\end{equation}
The usual way such plans are generated in the SE literature
is via some search-based method. Recently, we have made  some progress
in this area with the  GALE optimizer~\cite{krall14}. 
Like other search-based algorithms,
 GALE executes by proposing multiple $\delta$s
 (one for each generation in an evolutionary program).
 In practice, GALE makes
 extensive changes to initial sample of project data
 (perhaps 
100s to 100,000s of mutations). Each of these mutations
is reassessed using some domain model.
GALE uses an eigenvector technique to find better mutants faster than standard search-based SE tools~\cite{deb00a,zit02}.

When presented to some  software project managers, the GALE approach was challenged by
two business users. Note that the   issues they raise are
challenges not just  about GALE, but also  about any     optimization algorithm  that  
 make extensive use of mutation  and domain models.
One  user asked:
\begin{quote}
ISSUE~\#1: {\em ``What if I only want the minimum useful change, not
the maximum possible?''}
\end{quote}
Here, the business user is asking for solutions  that do not require extensive 
changes to their current  approaches.

@@

This paper addresses  ISSUE~\#1 and ISSUE~\#2 using case-based reasoning.
CBR  is purely based on the direct
adaptation of previous cases based on the similarity of
those cases with the current situation. That is, 
CBR based system have no explicit world model.
Rather, that model is implicit within  the available past
cases in the case cache.

This paper presents a novel CBR tool called ``HOW''. This tool  restricts test data $\delta$s to {\em minimal
displacements} (very small changes to a subset of the attributes)
thereby addressing ISSUE \#1. 
HOW ensures that the   $\delta$ displaces test data towards some region in the training data.
This is important since we address ISSUES~\#2 via some data mining
tool that train  a predictor from the training  data. Since we are pushing test data
into regions of the training data,  we can use the predictor to assess
the effectiveness of the $\delta$.
This paper shows that  that HOW's plans can be very effective indeed.
In the experiments shown below, HOW's plans
reduced the  expected
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
In summary, this paper makes five contributions:
\be
\item The HOW  case-based planner (which has not been presented before).
\item The experiments of this paper showing that HOW planner is useful for real-world tasks;
e.g. reducing the
expected values of the Jureczko software defects and the Seigmund software runtimes.
\item Apart from HOW, we  also offers a general framework for using  predictive system to build planners;
\item Further, we offer a  test to recognizes  when we should {\em not}  plan: specifically,
planning is
not recommended in domains with  poorly performing predictors;
\item Last, we  offer a general verification rig for assessing the value of plans generated within this rig.
\ee

\section{Background and Related Work}
\subsection{Search-based Software Engineering (SBSE)} CPUs are becoming
faster and cloud computing enables access to more CPUs. Hence, it is now pratical
to solve large problems in SBSE via 
genetic algorithms  that build multiple generations by  mutating  an existing population,  
the combine the best mutants to fond the next population~\cite{Harman2009,Harman2011}. Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}.

SBSE methods can make extensive mutations to the data they are exploring. Hence, these methods
require some domain model that can assess new and highly novel examples.

\subsection{Accessing Trustworthy Models}

In some domains, there is ready access to 
trustworthy domain model that can offer an assessment
of newly generated examples. For example, in other work
we have used executing source code as the ``model'' to check
if our mutations to test suites
  minimize
test suite size while maximizing the number of statements covered~\cite{me09m,andrews07,andrews10}.

However, not all domains come with ready-to-use models.
For   several problems where models are hard to obtain, or the input and output are related by complex connections that simply cannot be modeled in a reliable manner~\cite{Ludewig2003}. For example, 
consider software defect prediction and all the intricate issues that may lead defects in a product. A odel that includes {\em all} those
potential issues would be very large and unweliding. Further,
the empirical data required to validate any/all parts
of that model can be hard to find.

In fact, our experience has been that accessing and/or commissioning
a model can be a labor-intensive process.
For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).

Also, even when there is an existing model, they can require
constant  maintenance lest they become out-dated. Elsewhere, we have described our
extensions to the USC models to enable reasoning about agile software developments. 
It took many months to implement and certify those extensions~\cite{me09i,me09j}.
The problem of model maintenance is another
motivation to look for alternate methods that can be automatically updated whenever new data arrives.

Further, sometimes  it is easier to access cases of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

In summary, for domains with readily accessible models, we recommend
the kinds of tools that are widely used in the search-based
software engineering community such as GALE,NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like HOW.

\subsection{Planning and AI}
HOW is a {\em probablistic} planner, which is a different approach to that taken by
classical AI. 
That kind of planning is a logical procedure~\cite{Fikes1971}
that seeks an ordering on {\em operators} to take some domain
{\em state} from a {\em start} state to a {\em goal} state.
This classical logical approach is known to suffer from
computational bottlenecks~\cite{Bylander1994}. Hence,
like this paper,
many researchers take a probabilistic approach.
HOW's commitment to minimal displacement (and the fully-automated
modeling approach) of this paper makes it stand apart from, say,
  the Bayesian/Markov chain approach to planning of Dean et al.~\cite{dean1995planning}.

\subsection{Case-based Reasoning}
Case-based reasoning (CBR) has been extensively
explored in the AI literature
(where it has also bee called
 reasoning via anology or instance-based reasoning).
CBR seeks to emulate human recollection and adaptation
of past experiences in order to find solutions to current
problems.

CBR was first motivated by appeals to human cognition. According to 
Kolodner~\cite{Kolodner1992} and Shank~\cite{Schank1977},
humans do not base their decisions
on complex reductive analysis, but on an instantaneous survey of past experiences.
To say that   another way, according  to  CBR, 
{\em humans  don’t
think,  they remember.}

CBR has several advantages.
Firstly,
when the  cache of past cases is  updated and appended with additional cases, then a CBR system is instantly updated
to the latest findings.

Secondly, the output of a CBR is made on a ``case-by-case'' basis. 
That is, CBR does not offer some trite generalization over multiple
exmaples. Rather, the advice it gives is specialized to the particular
case being considered at the current time.  This is particularly attractive
since much of our recent work has focused on issues
of locality (where the best solutions for some part of the data
are very different some average case solution obtained from all the data~\cite{localvsglobal}).

\subsection{CBR in Software Engineering}
Due to the advantages of CBR, numerous papers in SE
have applied this technique.    CBR  has been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}. 
Researchers have found that, in some domains, they are competitive with  closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. 
CBR is also useful when dealing with spare data sets~\cite{1438374} of with 
partial knowledge  at, for example,  an early stage of a project)~\cite{walkerden1999empirical}.

 HOW    is very different
to the above CBR systems.
Firstly, HOW is a {\em planner} while the above
CBR methods are all {\em prediction} algorithms\footnote{The disctiniction being that
predictions systems tell you ``what is'', while planners tell you ``what to do''.}.
Secondly, all the above CBR tools explore the nearest ``k'' neighors,
HOW, on the other hand,  explores the {\em gradient between the scores of pairs of nearest neighbors}. 
To put that
another way,   the above CBR methods explore {\em points} while
HOW explores the slope {\em between} points.

\subsection{Prior Work}
HOW is the third
generation of our experiments with planning and CBR.
Generation one of that work resulted in
 W2, a CBR planner that reflected over the delta of raw  attributes~\cite{6600685}. It suffered from all-too-frequent {\em optimization failure}; i.e. when its plans were applied the resulting observed
performances {\em increased} in up to $\frac{1}{3}$rd of test cases.

Generation two of that work resulted in  IDEA~\cite{me12c}.
IDEA and HOW use the    same recursive clustering
method but IDEA made the mistake of  summarize those
recursive divisions into a tree structure. Initial results with IDEA were weak, but somewhat
positive~\cite{me12c}. We tried unsuccessfully for  months to extend and improve
the IDEA prototype. The results were most discouraging: those tree-based summaries often suffered
from the same optimization failure problem seen with W2.

As part of the IDEA work, we built HOW as a ``straw-man'' ; i.e. a  simpler
approach that was  to provide a baseline result,  above which IDEA was meant
to do better.   However, HOW's results were so good that we  threw away months
of work on tree-based planning with IDEA. Now, we strongly recommend HOW
over IDEA (and W2) since, as shown by the following results, HOW's plans {\em never} lead to performance getting worse. Also, when HOW did improve the expected values of the performance,
those performance improvements were an order of magnitude larger than those seen
with IDEA.

@@

\subsection{Design}
A general model for {\em predictors} is as follows. 
During testing, each test case gains a prediction 
by finding which region of the data it falls into.
During training, those regions are inferred from
divisions that ``chunk'' together regions with cases
with similar properties. These dividers can be categorical (as in the case of decision trees) or probabilistic (as in the
case of Naive Bayes) or learned via some distance minimization
process 
(as in the case of support vector machines) or generated
directly by some clustering algorithm.

@@

Data can be rapidly and recursively divided   as follows.
Find   two   distance cases,  $X,Y$
by picking any case $W$ at random, then setting $X$ to its most
distant case, then setting $Y$ to the case most distant from
$X$~\cite{fastmap}
(this requires only $O(2N)$ comparisons
of $N$ cases).
Next, HOW projects each case $Z$
onto a {\tt Slope} that  runs between $X,Y$ using the cosine
rule of \fig{where}b. After that,  split the data at the median $x$ value of all cases and
recurses on each half  (stopping when
one half has less  than $\sqrt{N}$ of the original population). For more details. \fig{howcode}, line~84,

@@

In the \fig{where}a, to compute distances, we use
the Euclidean measure recommended for
case-based reasoning by Aha et al.~\cite{aha91};
i.e. $\sqrt{\sum_iw_i(X_i-Y_i)^2}$ where $X_i,Y_i$
where values are  normalized 0..1 for the range min..max and 
$w_i$ is some importance weight given to attribute $i$.
Usually, $w_i=1$ but it can be adjusted using, say,
the feature weighting methods of \fig{where}d.

@@

Let   $X,Y$ be two points joined by  a {\em Slope} of  length $c$.
A third point, $Z$, has distances  $a=dist(Z,X)$ and
$b=dist(Z,Y)$ to $X,Y$, respectively.
According to the cosine rule,   $Z$ projects onto  $\overline{XY}$
at distance $x=(a^2 + c^2 - b^2)/(2c)$ from $X$.
Further, according to Pythagoras' theorem, $Z$ stands at a distance
$y = \sqrt{a^2 - x^2}$ from the line $\overline{XY}$. For more details, see  \fig{howcode}, line~52.

@@

HERE's feature weighting algorithm
comes from the CART regression tree learner~\cite{Breiman1984}.
It sorts independent variables
 according to how well they can reduce the variability
of a  numeric objective.
If we split some column $i$ of independent numeric data  into $N=n_1 + n_2 + ..$ splits,
then the expected
value of the objective value of the split  is $w_i = \sum_i v_in_i/N$
where $v$ is the standard deviation of the objective.
A recursive procedure can  find those divisions $n_1,n_2,...$ by (a)~sorting a numeric column,
then split at $\argmin_i w_i$, then recursing into each half.  For more details, see  \fig{howcode}, line~113.

@@

\input{tex/code}

Those same regions can be used for case-based 
{\em planning}. However,  instead of leaving test cases in the regions
found during training, we {\em displace} them towards regions
with better objective scores.
To achieve this, 
during training, HOW clusters data using the procedure of \fig{where}a,
then draws {\tt Slope}s between the centroids of  clusters $X,Y$ that are nearest
neighbors.
By convention, we say that the cluster containing $X$ has the  better objective scores. This means we can say that the {\tt Slope} is a useful direction for displacing
data; i.e.  away from $Y$ and towards $X$ (for an implementation of such a {\tt Slope} class, see \fig{howcode}, line~1).
 During testing,  HOW   finds the  nearest {\tt Slope} to every test case. That  test case is   then displaced towards the better end
of that {\tt Slope}; i.e. towards  $X$ (see \fig{howcode}, line 51).

 HOW needs to  determine the  exact {\em direction} and {\em magnitude} of 
displacement.  To address  ISSUE~\#1,
HOW uses two techniques to result in 
{\em minimal displacements} of project data. Firstly,
HOW sorts  the independent variables
using the procedure of  \fig{where}d. It then only
displaces the    best $B$\%  of that sort, while leaving the rest untouched
(see \fig{howcode}, line 38).
In  the experiments shown below,  we explore   \mbox{$B \in \{25,50,100\}\%$}.

Secondly, HOW   limits displacement
to very small regions in the space of the training data (never more that the separation
of the two closest clusters).  HOW's clusterer divides  data into many   small clusters containing
$\sqrt{N}$ of the population (using the {\tt cluster} function, \fig{howcode}, line 76).
HOW   limits displacement to  \mbox{$0 \le F \le c$}, where
$c$ is the distance between the two closest clusters. 
In  the experiments shown below,  we explore
\mbox{$F\in \{0.25, 0.5, 0.75\}$}.

Turning now to the displacement {\em direction},
our experiments show that it useful to  carefully reflect on exactly where HOW should  push each test case.
HOW replaces each cluster   with one {\em exemplar} that best
characterizes the data in that cluster. It then displaces the
test data towards that exemplar, as shown in \fig{howcode}, line 20. In summary, 
If  $Z_j$ is an independent variable within the
test case $Z$ and  $X$ is the exemplar found at training time,
then we displace $Z$ as follows:
\[Z_j =  Z_j + F*(X_j - Z_j)\]
Here, $F$ is the  magnitude controller, discussed above. Note that this
this equations satisfies our definition of a ``plan'' as offered in
\eq{one} in the  Introduction.

Different kinds of data require
different policies for exemplar generation.
For the Jureczko data, the objective score
is a boolean variable is true if defects have been noted
for that class.  For this kind of data (with discrete objectives),
we found nothing better than  making the exemplar of a cluster be the mean
value of all attributes of all cases in that cluster.

However, for the Seigmund data, we found a better exemplar generation policy.
\fig{pp} shows the effects of selecting different kinds of exemplars for 
Berkeley DB, one of the Seigmund data sets.
The \textcolor{black}{\bf black baseline} curve  shows the runtimes for 600 random configurations
of the Makefiles used in Berkely DB. The other two curves shows the predictions of the runtimes if test cases are displaced towards exemplars
  built from:
\bi
\item The \textcolor{red}{\bf   mean}    value of all cases in the  cluster.
\item The  case with the \textcolor{green}{\bf  best}  objective score. 
\ei
Note that using the best as the exemplar does better than useing the mean.
The pattern shown in \fig{pp} has been seen in all the Seigmund data sets;
i.e. for data sets with continuous classes,
best results come from using exemplars formed from the best instance (and not the mean).

@@

  The Seigmund data sets are far more varied than the Jurecko data.
  These contain the results of applying a large number of mutators
  to the configuration options of the Makefiles of  BerkeleyDB (the Java  and  Windows  versions),  Apache,  SQLite,  LLVM, and x264. For details on those datasets, see \fig{cpm}.
